Digging Deeper Into the State of the Practice for Domain Specific Research Software

### 2022-02-18 StateOfPractice/Papers/ICCS-2022-SOP_MethodologyPlusExamples/DigDeep.tex
Modified introduction section

The surveys used in previous studies have tended to recruit participants from
all domains of research software, possibly distinguishing them by programming
language (for instance, R developers~\cite{PintoEtAl2018}), or by the role of
the developers (for instance postdoctoral researchers \cite{UditAndKatz2017}).
Case studies, on the other hand, have focused on a few specific examples at a
time, as is the nature of case studies.  For our new methodology, we propose a
scope between these two extremes.  Rather than focus on all research software,
or just a few examples, we will focus on one domain of research software at a
time. The practical reason for this scope is that digging deep into
repository data takes time, making a broad scope infeasible. We have imposed
a practical constraint of one person month worth of work per domain.\footnote{A person
month is considered to be $20$ working days ($4$ weeks in a month, with $5$ days
of work per week) at $8$ person hours per day, or $20 \cdot 8 = 160$ person
hours.} Focusing on one domain at a time has more than just practical
advantages.  By restricting ourselves to a single domain we can bring domain
knowledge and domain experts into the mix.  The domain customized insight
provided by the assessment has the potential to help a specific domain as they
adopt and develop new development practices.  Moreover, measuring different
domains facilitates uncovering domain specific practices, which can later be
compared and contrasted.

@@

Our methodology answers the research question through inspecting repositories,
using the Analytic Hierarch Process (AHP) for ranking software, interviewing
developers and interacting with at least one domain expert.  We leave the
measurement of the performance, for instance using benchmarks, to other projects
\cite{KaagstromEtAl1998}. The current methodology updates the approach used in
prior assessments of domains like Geographic Information Systems
\cite{SmithEtAl2018_arXivGIS}, Mesh Generators \cite{SmithEtAl2016}, Seismology
software \cite{SmithEtAl2018}, and statistical software for psychology
\cite{SmithEtAl2018_StatSoft}.  Initial tests of the new methodology have been
done for medical image analysis software \cite{Dong2021} and for Lattice
Boltzmann Method (LBM) software \cite{Michalski2021}.  The LBM example will be
used throughout this paper to illustrate the steps in the methodology.  The
paper ends with a summary of potential threats to validity
(Section~\ref{threats}) and our final conclusions and recommendations
(Section~\ref{SecConcludingRemarks}).

>>>>

The surveys used in previous studies have tended to recruit participants from
all domains of research software, possibly distinguishing them by programming
language (for instance, R developers~\cite{PintoEtAl2018}), or by the role of
the developers (for instance postdoctoral researchers \cite{UditAndKatz2017}).
Case studies, on the other hand, have focused on a few specific examples at a
time, as is the nature of case studies.  For our new methodology, we propose a
scope between these two extremes.  Rather than focus on all research software,
or just a few examples, we will focus on one domain at a time. The practical
reason for this scope is that digging deep takes time, making a broad scope
infeasible. We have imposed a practical constraint of one person month of effort
per domain.\footnote{A person month is considered to be $20$ working days ($4$
weeks in a month, with $5$ days of work per week) at $8$ person hours per day,
or $20 \cdot 8 = 160$ person hours.} Focusing on one domain at a time has more
than just practical advantages.  By restricting ourselves to a single domain we
can bring domain knowledge and domain experts into the mix.  The domain
customized insight provided by the assessment has the potential to help a
specific domain as they adopt and develop new practices.  Moreover, measuring
multiple different domains facilitates comparing and contrasting domain specific
practices.

@@

Our methodology answers the research question through inspecting repositories,
using the Analytic Hierarch Process (AHP) for ranking software, interviewing
developers and interacting with at least one domain expert.  We leave the
measurement of the performance, for instance using benchmarks, to other projects
\cite{KaagstromEtAl1998}. The current methodology updates the approach used in
prior assessments of domains like Geographic Information Systems
\cite{SmithEtAl2018_arXivGIS}, Mesh Generators \cite{SmithEtAl2016}, Seismology
software \cite{SmithEtAl2018}, and statistical software for psychology
\cite{SmithEtAl2018_StatSoft}.  Initial tests of the new methodology have been
done for medical image analysis software \cite{Dong2021} and for Lattice
Boltzmann Method (LBM) software \cite{Michalski2021}.  The LBM example will be
used throughout this paper to illustrate the steps in the methodology.  The
paper ends with a summary of potential threats to validity
(Section~\ref{threats}) and our conclusions
(Section~\ref{SecConcludingRemarks}).
